### 2014-11-24 paper/draft.tex

\documentclass[runningheads,a4paper]{llncs}

\usepackage{amssymb}
\setcounter{tocdepth}{3}
\usepackage{graphicx}
\usepackage{amsmath}
%\usepackage{amsfonts}
%\usepackage{amsthm}
\usepackage{subfigure}
%\usepackage{caption} 
%\usepackage{subcaption}
%\usepackage{cite}
\usepackage{hyperref}
\usepackage{url}
\urlstyle{same}
\newcommand{\keywords}[1]{\par\addvspace\baselineskip
\noindent\keywordname\enspace\ignorespaces#1}

\makeatletter
\let\c@lemma=\c@theorem
\let\c@corollary=\c@theorem
\let\c@fact=\c@theorem
\makeatother

\let\realendproof=\endproof
\def\endproof{\hspace*{\fill}$\Box$\realendproof}

\begin{document}

\mainmatter

\title{Experiments on ANTS Algorithms Without Communication}
\author{Fermi Ma \and Nicolas Rakover \and Erik Waingarten}
\institute{MIT}

\maketitle

\section{Introduction}

\subsection{Related Works}

\section{Algorithms}

We run experiments on three algorithms:
\begin{itemize}
\item FKLS
\item LLNR
\item Harmonic search algorithm
\end{itemize}

Go through thorough theoretical analysis of each. Explain lower bound. Give the model and the motivations.

\subsection{Extending LLNR to uniform}

\section{Experiments and Results}

\subsection{Non-uniform in $n$}

\subsection{Uniform in $n$}

\section{Conclusion}

\bibliographystyle{plain}
\bibliography{references}

\end{document}